##### see https://learnr.wordpress.com/2009/03/29/ggplot2_marimekko_mosaic_chart/
##### note that this works with the "opts" part commented out (opts is deprecated in ggplot2)
##### another note: this file looks horrible in Github, try knitting it as an Rmd in Rstudio
---
title: "Marimekko"
author: "Eric, hastily copying and pasting from
  https://learnr.wordpress.com/2009/03/29/ggplot2_marimekko_mosaic_chart/
  in case that website is taken down before he gets around to writing a function
  that does this"
output: html_document
---


```{r}
library(reshape)
library(plyr)
library(ggplot2)
```

We start with creating the dataframe to work with…
```{r}
df <- data.frame(segment = c("A", "B", "C",
                               "D"), segpct = c(40, 30, 20, 10), Alpha = c(60,
                                                                           40, 30, 25), Beta = c(25, 30, 30, 25),
                   Gamma = c(10, 20, 20, 25), Delta = c(5,
                                                        10, 20, 25))
```
Next, I add a few helper variables, and convert the data to long format.

Calculate cumulative width on x-axes, starting point of each column and erase segpct variable.
```{r}
df$xmax <- cumsum(df$segpct)
df$xmin <- df$xmax - df$segpct
df$segpct <- NULL
```
Data looks like this before the long-format conversion:
```{r}
head(df)
```
After melting the data looks like this:
```{r}  
library(ggplot2)
dfm <- melt(df, id = c("segment", "xmin", "xmax"))
head(dfm)
```
Now we need to determine how the columns are stacked and where to position the text labels.

Calculate ymin and ymax:
```{r}  
dfm1 <- ddply(dfm, .(segment), transform, ymax = cumsum(value))
dfm1 <- ddply(dfm1, .(segment), transform,
                ymin = ymax - value)
```
Positioning of text:
```{r}  
dfm1$xtext <- with(dfm1, xmin + (xmax - xmin)/2)
dfm1$ytext <- with(dfm1, ymin + (ymax - ymin)/2)
```
Finally, we are ready to start the plotting process:
```{r}
p <- ggplot(dfm1, aes(ymin = ymin, ymax = ymax,
                          xmin = xmin, xmax = xmax, fill = variable))
```
Use grey border to distinguish between the segments:
```{r}  
p1 <- p + geom_rect(colour = I("grey"))
```

The explanation of different fill colours will be included in the text label of Segment A using the ifelse function.
```{r}
p2 <- p1 + geom_text(aes(x = xtext, y = ytext,
                           label = ifelse(segment == "A", paste(variable,
                                                                " - ", value, "%", sep = ""), paste(value,
                                                                                                    "%", sep = ""))), size = 3.5)
```
The maximum y-axes value is 100 (as in 100%), and to add the segment description above each column I manually specify the text position.
```{r}
p3 <- p2 + geom_text(aes(x = xtext, y = 103,
                           label = paste("Seg ", segment)), size = 4)
```
Some last-minute changes to the default formatting: remove axis labels, legend and gridlines.
```{r}
p3 + theme_bw() + labs(x = NULL, y = NULL,
                         fill = NULL) 
#+ opts(legend.position = "none",
#                                             panel.grid.major = theme_line(colour = NA),
#                                             panel.grid.minor = theme_line(colour = NA))
```
Or, if the default palette is not to one’s liking it is very easy to use a ColorBrewer palette instead:
```{r}
last_plot() + scale_fill_brewer(palette = "Set2")
```{r}
